---
title: "Practical Machine Learning"
author: "Noel Namai"
date: "Sunday, July 20, 2014"
output: html_document
---

### Data Collection:

First a function to read data from respective CSV file. It keeps the header and replaces all "NA", "" and "#DIV/0!" with "NA"

```{r}
readData <- function(file) {
    df <- read.csv(file,
                   header=TRUE, 
                   na.strings=c("NA", "", "#DIV/0!"), 
                   sep=",")
    }
```

Then the data is read using the read **readData** function:

```{r}
data <- readData("./data/training.csv") #read data
```

### Data Cleaning:

Then the function **cleanData** will be used to clean data. This function turns the variable **new_window** from "yes" or "no" into "1" or "0" respectively. It converts the variable **cvtd_timestamp** into a time object and splits it into **year, month, weekday, hour** and **minute** variables.

Then all the columns with the most "NA" are droped from the dataframe. Columns **X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp** are also droped.

```{r}
cleanData <- function(df) {    
    df$new_window <- ifelse(df$new_window=="yes", "1", "0") #convert new_window into "0" or "1"
    df$cvtd_timestamp <- strptime(df$cvtd_timestamp, "%d/%m/%Y %H:%M") #convert cvtd_timestamp into a time object
    df$year <- as.numeric(strftime(df$cvtd_timestamp, "%Y")) #creat a new feature year
    df$month <- as.numeric(strftime(df$cvtd_timestamp, "%m")) #creat a new feature month
    df$weekday <- as.numeric(strftime(df$cvtd_timestamp, "%d")) #creat a new feature weekday
    df$hour <- as.numeric(strftime(df$cvtd_timestamp, "%H")) #creat a new feature hour
    df$minute <- as.numeric(strftime(df$cvtd_timestamp, "%M")) #creat a new feature minute
    df <- df[,colSums(is.na(df)) < 1] #drop columns with the most "NA"
    df <- subset(df, select=-c(X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp)) #drop noisy features 
    }
```

Then the data is cleaned using the **cleanData** function:

```{r}
data <- cleanData(data) #clean data
```

### Training:

The function **createDataPartition** can be used to create a stratified random sample of the data into **training** and **validation** sets with **70%** of the data in the **training** set and the rest in the **validation** set:

```{r, warning=FALSE, message=FALSE}
library(caret)
set.seed(1)
inTrain <- createDataPartition(y=data$classe, p=0.7, list=FALSE)
training <- data[inTrain,] #create training set
testing <- data[-inTrain,] #create validation set
```

The model is fitted using **train** from the **caret** library with the following parameters:

* **method** : argument specifies the type of training model.

* **tuneGrid** : a data frame with columns for each tuning parameter.

* **trControl** :  used to specifiy the type of resampling.

    * **method** : the resampling method to be used.

    * **number** : number of cross-validation groups. This may also be an explicit list of integers that define the cross-validation groups.

```{r train_model, warning=FALSE, message=FALSE}
#creat a tune grid
grid <-  expand.grid(cp=c(1:10)*0.01)

#fit the classification model
fit <- train(classe ~ ., data=training,
             method="rpart",
             tuneGrid=grid,
             trControl=trainControl(method="cv", number=10))

fit
plot(fit)
```

```{r, warning=FALSE, message=FALSE, fig.width=10, fig.height=5}
library(rattle)
fancyRpartPlot(fit$finalModel)
```
